Dynamic Transition Analysis with TIMES

### 2020-06-21 doc/2020-03-paper/introduction.tex
updates to section 3
Former-commit-id: 75b8a4e4d2f916c3fbd93ac61343e53675edb58a
Former-commit-id: 89fa8262a5e8ce794da5c196f7937e9ad3251205

\section{Introduction} \label{Introduction}
In order to mitigate climate change, and to improve environmental outcomes, many nations are actively seeking to reduce carbon emissions, and have formalized this goal through the Paris Agreements (UNFCCC, 2015). The largest contribution to global greenhouse gas (GHG) emissions, some 73\%, comes from energy consumption, consisting mainly of the transportation, electricity and heat, buildings, manufacturing and construction sectors (WRI, 2020). For developed economy nations rich in natural resources, switching to natural gas or implementing carbon capture and storage (CCS) on fossil fuel plants is one option to reduce GHG emissions, however for developing nations, these options are not always economically feasible, leading to increased emissions (through greater coal use) for rapidly developing nations (IEA, 2019). For Japan, a developed economy nation with no fossil fuel resources, the challenge to reduce GHG emissions is likely to follow a different path, already evidenced by the restart of nuclear reactors and a shift toward large scale renewable energy deployment (IEA, 2019).
Although influenced by the Paris Agreements, Japanese energy policy is governed by the Basic Energy Plan, recently updated to the 5th edition and approved by the Cabinet in July of 2018 (METI, 2018). The Basic Energy Plan outlines national policy toward a new energy system for the years 2030 and 2050, cognizant of limited indigenous resources, the impact of the Fukushima incident and external pressures on energy supplies (ANRE, 2018). The Plan reaffirms the Japanese benchmarks for evaluating the energy system as, first and foremost within the context of energy security, followed by economic efficiency, safety, and a consideration of the environment (summarized as '3E+S'; ibid). Although there are some parallels between the 3E+S goals and the Paris Agreement targets agreed to by Japan, the plan does not set out in detail how the 2050 emission reduction target of 80\% is to be met. This may be due to the lack of evidence for the economically feasible reduction of GHG emissions in each sector, and it has been suggested that electrification of a number of sectors will be required to achieve the ambitious 2050 target, underpinned by electricity generated from low-carbon technologies (Matsuo et al., 2018). For the power sector to achieve such a target, near-zero emissions are required, and early action utilizing existing technologies is preferable to delaying action in preference for future technologies (Ashina et al., 2012). It is likely that a mixture of current and emerging technologies will be employed to achieve carbon reduction targets in Japan. The current candidates include a reinvigoration of the nuclear contribution to final energy demand, deploying CCS to fossil fuel power plants, and the ushering in of the hydrogen economy, underpinned by renewable energy deployment as well as hydrogen imports from abroad (Ashina et al., 2012; Matsuo et al., 2018; METI, 2017). 
The aim of this research is to investigate the likely suite of technologies and their feasibility in meeting Japan's carbon reduction goals, cognizant of energy policy, resource limitations, demand growth, emerging technologies and economic constraints using the TIMES framework. Our dynamic simulations of transition scenarios, by focusing on minimizing the cost of the transition while satisfying CO2 emission constraints, suggest potential economically feasible decarbonization pathways while meeting the increasing near-term electricity demand. Additionally, the significance of key economic parameters of emerging technologies is assessed through sensitivity analysis, in order to highlight the most impactful parameters of each technology and hence guide research and development efforts focused on these technologies.

>>>>

\section{Introduction} \label{Introduction}
In order to mitigate climate change, and to improve environmental outcomes, many nations are actively seeking to reduce carbon emissions, and have formalized this goal through the Paris Agreements (UNFCCC, 2015). The largest contribution to global greenhouse gas (GHG) emissions, some 73\%, comes from energy consumption, consisting mainly of the transportation, electricity and heat, buildings, manufacturing and construction sectors (WRI, 2020). For developed economy nations rich in natural resources, switching to natural gas or implementing carbon capture and storage (CCS) on fossil fuel plants is one option to reduce GHG emissions, however for developing nations, these options are not always economically feasible, leading to increased emissions (through greater coal use) for rapidly developing nations (IEA, 2019). For Japan, a developed economy nation with no fossil fuel resources, the challenge to reduce GHG emissions is likely to follow a different path, already evidenced by the restart of nuclear reactors and a shift toward large scale renewable energy deployment (IEA, 2019).

Although influenced by the Paris Agreements, Japanese energy policy is governed by the Basic Energy Plan, recently updated to the 5th edition and approved by the Cabinet in July of 2018 (METI, 2018). The Basic Energy Plan outlines national policy toward a new energy system for the years 2030 and 2050, cognizant of limited indigenous resources, the impact of the Fukushima incident and external pressures on energy supplies (ANRE, 2018). The Plan reaffirms the Japanese benchmarks for evaluating the energy system as, first and foremost within the context of energy security, followed by economic efficiency, safety, and a consideration of the environment (summarized as '3E+S'; ibid). Although there are some parallels between the 3E+S goals and the Paris Agreement targets agreed to by Japan, the plan does not set out in detail how the 2050 emission reduction target of 80\% is to be met. This may be due to the lack of evidence for the economically feasible reduction of GHG emissions in each sector, and it has been suggested that electrification of a number of sectors will be required to achieve the ambitious 2050 target, underpinned by electricity generated from low-carbon technologies (Matsuo et al., 2018). For the power sector to achieve such a target, near-zero emissions are required, and early action utilizing existing technologies is preferable to delaying action in preference for future technologies (Ashina et al., 2012). It is likely that a mixture of current and emerging technologies will be employed to achieve carbon reduction targets in Japan. The current candidates include a reinvigoration of the nuclear contribution to final energy demand, deploying CCS to fossil fuel power plants, and the ushering in of the hydrogen economy, underpinned by renewable energy deployment as well as hydrogen imports from abroad (Ashina et al., 2012; Matsuo et al., 2018; METI, 2017).

The aim of this research is to investigate the likely suite of electricity generation and storage technologies and their feasibility in meeting Japan's carbon reduction goals, cognizant of energy policy, resource limitations, demand growth, emerging technologies and economic constraints using the TIMES framework. Our dynamic simulations of transition scenarios, by focusing on minimizing the cost of the transition while satisfying CO2 emission constraints, suggest potential economically feasible decarbonization pathways while meeting the increasing near-term electricity demand. Additionally, the significance of key economic parameters of emerging technologies is assessed through sensitivity analysis, in order to highlight the most impactful parameters of each technology and hence guide research and development efforts focused on these technologies.

### 2020-06-21 doc/2020-03-paper/litreview.tex
updates to section 3
Former-commit-id: 75b8a4e4d2f916c3fbd93ac61343e53675edb58a
Former-commit-id: 89fa8262a5e8ce794da5c196f7937e9ad3251205

The Paris Agreement commits individual nations to significant carbon reduction over time through the Intended Nationally Determined Contribution (INDC) mechanism (UNFCCC, 2019). Japan, as a signatory to the Paris Agreements has submitted an INDC with the following goals and timelines: by 2030, to reduce GHG emissions by 26\% compared to 2013 levels, and by 2050 to reduce overall GHG emissions by 80\% or more, through the “development and diffusion of low-carbon technologies and transition to a low-carbon socio-economic structure” (UNFCCC, 2015). Cognizant of these targets, a number of authors have evaluated Japan’s future energy system using a variety of modeling approaches. Using the MARKAL model, considering the uncertainties of technology development, Ozawa et al., found that hydrogen will play a major role in the future energy system, reliant on both nuclear power and carbon capture and storage (CCS) to reduce electricity sector emissions to nearly zero by 2050 (2018). Recognizing the benefits that renewable energy will play in reducing carbon emissions, and the issues of intermittency, Li et al. explored the role of hydrogen as a storage medium through power to gas (P2G) approaches in Kyushu, Japan. Their study identified that P2G can increase the effective utilization rate of renewable energy, and the use of hydrogen in the gas network, effectively pairing the electricity and gas networks, overcomes current renewable electricity curtailment issues (2019). Cognizant of the Japanese government’s strategic approach to carbon reduction out to 2050 via energy system reform, Chapman & Pambudi also identify a strong role for nuclear, renewables and hydrogen under a carbon constrained, optimal cost MARKAL/TIMES simulation approach (2018). Considering economic conditions and demographic trends such as moderate GDP growth and rapid aging which are occurring in Japan, Kuriyama et al., suggest that 2030 targets can be met or exceeded (i.e. up to 42\% GHG reduction) with limited renewable energy growth at a 15\% contribution from nuclear, or even without nuclear, under a renewable growth scenario (2019). It appears that these trends and energy system changes will be insufficient to meet the more ambitious 2050 targets. Taking a more holistic view, in line with the Japanese government’s 3E+S targets, ambitious research and development to enable high levels of renewable deployment appears to be necessary to not only meet deep emission reduction goals, but to also reduce Japanese dependence on imported fuels, cognizant of both CCS and nuclear deployment rates into the future. Consensus on policy options and priorities also have a large influence on modeling outcomes for the  Low Carbon Navigator assessing Japanese energy and emission options out to 2050 (Moinuddin & Kuriyama, 2019). A seminal work by Sugiyama et al., brings together a number of modeling approaches for Japan’s long term (i.e. to the year 2050) climate change mitigation options, utilizing national and global general and partial equilibrium models (2019). Model results are contrasted under six scenarios which incorporate a baseline and a range of emissions reductions (50-80\%) and regional obligations for global models (ibid.). under the Paris Agreement target of an 80\% reduction, each of the models assessed recognize the importance of renewable energy deployment by 2050, notably hydro, solar and wind, with varying contributions from nuclear energy and fossil fuels, predominantly natural gas. Additionally, for Japan, the option for import of carbon-free hydrogen was identified as potentially playing a critical role (Sato, 2005; Akimoto et al., 2010; Matsuo et al., 2013; Fujii and Komiyama, 2015; Oshiro & Masui, 2015; Sugiyama et al., 2019). Many studies consider hydrogen a critical part of Japan’s low-carbon energy transition, specifically toward improving energy security, its ability to be produced from multiple sources and lack of emissions when combusted (Iida & Sakata, 2019). Global modelling efforts consider the incorporation of long-distance international transport of hydrogen with end uses dominated by fuel cell passenger and freight vehicles and power generation, via both mixed and direct combustion. Electricity from hydrogen is estimated to emerge in Japan from 2030 onwards, as nuclear and coal fired power generation reduce toward 2050 (Ishimoto et al., 2017). From a policy standpoint, Japan has committed to achieving a hydrogen society with the primary goal of cost parity of hydrogen with competing fuels, requiring a three-fold reduction in cost by 2030, and further reductions into the future (Nagashima, 2018). Under the Basic Hydrogen Strategy, the Japanese government aims to realize low cost hydrogen use in power generation, mobility and industry, develop international supply chains to ensure stable supply, expand renewable deployment and revitalize regional areas and develop hydrogen related technologies (METI, 2017). The strategy aims to account for both economic and geopolitical impacts and the need to prioritize research and development to overcome the economic and technical challenges (Nagashima, 2018). A common thread across previous research is that the uncertainty surrounding carbon capture and storage (particularly with regard to scaling up and public acceptance issues) and the role that nuclear energy will play, largely due to policy reform which occurred in response to the nuclear accident in 2011 (Oshiro et al., 2019).

This work leverages the dynamic simulation capabilities of TIMES <CITE LOULOU> by incorporating learning curves for parameters such as investment and o&m costs, efficiency, and emission coefficients. Our model also incorporates life-cycle emissions of all conventional and emerging technologies. This allows for a more meaningful analysis of the global warming potential of emerging technologies, as life-cycle emissions become significant when considering deep emission reductions of the order of 80\%. The model and approach proposed in this research is unique, as it builds on the works and modeling consensus outlined in the literature review and expands the consideration of technologies beyond conventional and recent technologies to also include emerging and disruptive technologies post-2050. By modeling the Japanese energy system out to the year 2100, our aim is to detail the mid and long-term impacts of technological development and market penetration, and to identify the suite of technologies which could underpin the successful achievement of carbon reduction.

>>>>

The Paris Agreement commits individual nations to significant carbon reduction over time through the Intended Nationally Determined Contribution (INDC) mechanism (UNFCCC, 2019). Japan, as a signatory to the Paris Agreements has submitted an INDC with the following goals and timelines: by 2030, to reduce GHG emissions by 26\% compared to 2013 levels, and by 2050 to reduce overall GHG emissions by 80\% or more, through the "development and diffusion of low-carbon technologies and transition to a low-carbon socio-economic structure" (UNFCCC, 2015). Cognizant of these targets, a number of authors have evaluated Japan's future energy system using a variety of modeling approaches. Using the MARKAL model, considering the uncertainties of technology development, Ozawa et al., found that hydrogen will play a major role in the future energy system, reliant on both nuclear power and carbon capture and storage (CCS) to reduce electricity sector emissions to nearly zero by 2050 (2018). Recognizing the benefits that renewable energy will play in reducing carbon emissions, and the issues of intermittency, Li et al. explored the role of hydrogen as a storage medium through power to gas (P2G) approaches in Kyushu, Japan. Their study identified that P2G can increase the effective utilization rate of renewable energy, and the use of hydrogen in the gas network, effectively pairing the electricity and gas networks, overcomes current renewable electricity curtailment issues (2019). Cognizant of the Japanese government's strategic approach to carbon reduction out to 2050 via energy system reform, Chapman and Pambudi also identify a strong role for nuclear, renewables and hydrogen under a carbon constrained, optimal cost MARKAL/TIMES simulation approach (2018). Considering economic conditions and demographic trends such as moderate GDP growth and rapid aging which are occurring in Japan, Kuriyama et al., suggest that 2030 targets can be met or exceeded (i.e. up to 42\% GHG reduction) with limited renewable energy growth at a 15\% contribution from nuclear, or even without nuclear, under a renewable growth scenario (2019). It appears that these trends and energy system changes will be insufficient to meet the more ambitious 2050 targets. Taking a more holistic view, in line with the Japanese government's 3E+S targets, ambitious research and development to enable high levels of renewable deployment appears to be necessary to not only meet deep emission reduction goals, but to also reduce Japanese dependence on imported fuels, cognizant of both CCS and nuclear deployment rates into the future. Consensus on policy options and priorities also have a large influence on modeling outcomes for the  Low Carbon Navigator assessing Japanese energy and emission options out to 2050 (Moinuddin \& Kuriyama, 2019). A seminal work by Sugiyama et al., brings together a number of modeling approaches for Japan's long term (i.e. to the year 2050) climate change mitigation options, utilizing national and global general and partial equilibrium models (2019). Model results are contrasted under six scenarios which incorporate a baseline and a range of emissions reductions (50-80\%) and regional obligations for global models (ibid.). under the Paris Agreement target of an 80\% reduction, each of the models assessed recognize the importance of renewable energy deployment by 2050, notably hydro, solar and wind, with varying contributions from nuclear energy and fossil fuels, predominantly natural gas. Additionally, for Japan, the option for import of carbon-free hydrogen was identified as potentially playing a critical role (Sato, 2005; Akimoto et al., 2010; Matsuo et al., 2013; Fujii and Komiyama, 2015; Oshiro \& Masui, 2015; Sugiyama et al., 2019). Many studies consider hydrogen a critical part of Japan's low-carbon energy transition, specifically toward improving energy security, its ability to be produced from multiple sources and lack of emissions when combusted (Iida \& Sakata, 2019). Global modelling efforts consider the incorporation of long-distance international transport of hydrogen with end uses dominated by fuel cell passenger and freight vehicles and power generation, via both mixed and direct combustion. Electricity from hydrogen is estimated to emerge in Japan from 2030 onwards, as nuclear and coal fired power generation reduce toward 2050 (Ishimoto et al., 2017). From a policy standpoint, Japan has committed to achieving a hydrogen society with the primary goal of cost parity of hydrogen with competing fuels, requiring a three-fold reduction in cost by 2030, and further reductions into the future (Nagashima, 2018). Under the Basic Hydrogen Strategy, the Japanese government aims to realize low cost hydrogen use in power generation, mobility and industry, develop international supply chains to ensure stable supply, expand renewable deployment and revitalize regional areas and develop hydrogen related technologies (METI, 2017). The strategy aims to account for both economic and geopolitical impacts and the need to prioritize research and development to overcome the economic and technical challenges (Nagashima, 2018). A common thread across previous research is that the uncertainty surrounding carbon capture and storage (particularly with regard to scaling up and public acceptance issues) and the role that nuclear energy will play, largely due to policy reform which occurred in response to the nuclear accident in 2011 (Oshiro et al., 2019).

This work leverages the dynamic simulation capabilities of TIMES (Loulou et al., 2005) by incorporating learning curves for parameters such as investment and \gls{OM} costs, efficiency, and emission coefficients. Our model also incorporates life-cycle emissions of all conventional and emerging technologies. This allows for a more meaningful analysis of the global warming potential of emerging technologies, as life-cycle emissions become significant when considering deep emission reductions of the order of 80\%. The model and approach proposed in this research is unique, as it builds on the works and modeling consensus outlined in the literature review and expands the consideration of technologies beyond conventional and recent technologies to also include emerging and disruptive technologies post-2050. By modeling the Japanese electricity supply system out to the year 2100, our aim is to detail the mid and long-term impacts of technological development and market penetration, and to identify the suite of technologies which could underpin the successful achievement of carbon reduction.

### 2020-06-21 doc/2020-03-paper/methods.tex
updates to section 3
Former-commit-id: 75b8a4e4d2f916c3fbd93ac61343e53675edb58a
Former-commit-id: 89fa8262a5e8ce794da5c196f7937e9ad3251205

\section{Model description} \label{Model-description}

\subsection{TIMES Model Description}
Objective function.
Constraints - CO2 emissions reduction targets, increasing demand, CO2 tax.
Important parameters.

\subsection{Common Features - Initial condition and constraints}

>>>>

\section{Methodology} \label{method}

\subsection{TIMES Model Description}
The \gls{TIMES} model generator is designed to model dynamic energy systems and simulate transition scenarios as a mixed-integer programming linear optimization problem that is subject to a primary objective function and additional constraints. The generation, trade, refinement, storage, and supply of energy commodities across multiple sectors and multiple regions is modeled using a wide variety of in-built commodity and process types. Emissions can be associated with energy commodities or processes as emission coefficient per unit commodity produced or consumed.

The objective function in our simulations is the overall cost of the transition, hence the primary objective of the solver is to minimize said cost.The major constrains in our simulations are the overall demand for electricity, emission constraints based on Japan's \gls{NDC}, and feasible nameplate capacity deployment limits.
